---
title: "Current Structure of the Mexican Energy System"
author: "Raul Gutierrez and Luis Sarmiento"
format: 
  html: 
    toc: true
    warning: false
    code-fold: true
    code-summary: "Show the Code"
---

# Introduction

We divide the Mexican energy system into three main sectors; power, transport, and high-temperature heating.^[We do not include low-temperature (or residential) heating because the share of energy demand for heating in Mexico is marginal and the energy demand for cooling is connected to the power-sector demand for electricity].

```{r}
#| echo: false
conflict_prefer("week", "lubridate")
conflict_prefer("filter", "dplyr")
conflict_prefer("isoweek", "lubridate")
conflict_prefer("map", "purrr")
conflict_prefer("year", "lubridate")
conflict_prefer("month", "lubridate")
conflict_prefer("select", "dplyr")
conflict_prefer("lag", "dplyr")
```

```{r}
#| echo: false
# Change the working directory
file = gsub("WebSite-MeritOrder", "", getwd())
# Load the demand file
data = read_csv(paste0(file, "00_RawData/modelData/general/par_demandRest.csv"), show_col_types = F)
# Select relevant variables
data = data |> select(region = region_2, sector = carrier_1, year = timestep_1,
                      TimeStep = timestep_2, value)
# Capitalize and standardize the name of the regions
data = mutate(data, region = str_to_title(gsub("([a-z])([A-Z])", "\\1 \\2", region)))
```

@fig-SectorDemand contains the exogenous energy demand across sectors we use for the optimization. We use data from the national power development program to project the demand forward. The sector with the highest energy demand is high temperature (industrial) heating, closely followed by power generation. At the start of the optimization period (2020), high-temperature heating consumes 40% of total energy demand, the electric sector 38%, and passenger transportation 21%. @fig-Map shows the map of the national center for energy control (CENACE) power system regions. There are nine regions; Baja California, Baja California Sur, Noroeste, Norte, Noreste, Occidental, Central, Oriental, and Peninsula. @fig-RegionalPowerDemand portrays the exogenous energy demand for the power sector between 2020 and 2050 across these regions. While in 2020, there was an aggregate demand of 567 PJ, by 2050, the market will grow by 50% to more than 1,100 PJ. Regarding the distribution of power demand across all regions. The peninsular region uses the least amount of electricity with only 4% of total resources, while the industrious Western, Central, and Northeast areas demand 22%, 18%, and 17% of power supply, respectively. 

::: panel-tabset
## Demand by Sector
```{r}
#| label: fig-SectorDemand 
#| fig-cap: Demand for energy across sectors in PetaJoules
# Change the working directory
plot = data |> group_by(year, sector) |> 
  summarise(demand = sum(value, na.rm = T)) |> 
  dplyr::filter(sector != "heatLow", year > 2015) |> group_by(year) |>  
  mutate(share = demand/sum(demand))
# Plot the power demand
ggplot(plot) +
  geom_line(aes(x = year, y = demand, group = sector, color = sector)) +
  geom_point(aes(x = year, y = demand, group = sector, color = sector))  +
  theme_economist()  


```
## Regions
```{r}
#| label: fig-Map
#| fig-cap: Demand for energy across sectors in PetaJoules
# Change the working directory
file = gsub("WebSite-MeritOrder", "", getwd())
# Load the shapefile of states
st = st_read(paste0(file, "/00_RawData/99_other/01_shp/01_mx/gadm40_MEX_1.shp"), quiet = T)
reg = read_delim(paste0(file, "/00_RawData/99_other/99_other/MxRegions.csv"), 
    delim = ";", escape_double = F, trim_ws = T)
# Change the state names to merge with the region codes
st = mutate(st, NAME_1 = janitor::make_clean_names(NAME_1)) |> 
  mutate(NAME_1 = gsub("_", " ", NAME_1)) |>  mutate(NAME_1 = str_to_title(NAME_1))
# Change the working directory
plot = left_join(st, reg, by = c("NAME_1" = "state"))
# Plot the power demand
tmap_style("classic"); tm_graticules(labels.inside.frame = T, labels.show = F) +
  tm_shape(plot) + tm_fill(col = "region", title = "")  +
  tm_compass(type = "rose", position = c("left", "top"), size = 2) +
  tm_scale_bar(width = 0.2, position = c("left", "bottom"), text.size = 1)


```
## Power
```{r}
#| label: fig-RegionalPowerDemand 
#| fig-cap: Energy demand in the power sector across regions
# Change the working directory
plot = data |> dplyr::filter(sector == "power") |> 
  mutate(region = gsub("Baja California Sur", "Baja California", region)) |> 
  group_by(year, region) |> 
  summarise(demand = sum(value, na.rm = T)) |> 
  dplyr::filter( year > 2015) |> arrange(year, -demand) |> 
  mutate(region = fct_reorder(factor(region), demand, .desc = F)) |>
  group_by(year) |> 
  mutate(share = paste0("(", as.character(scales::comma(round(demand/sum(demand)*100, 0))), "%", ")")) |> 
  mutate(region = paste(region, share)) |> group_by(year) |> mutate(agg = sum(demand))
head(plot)
# Plot the power demand
ggplot(plot) +
  geom_bar(aes(x = year, y = demand, fill = reorder(region, demand)), alpha = 0.95, stat = "identity") +
  theme_economist() %+replace% theme(legend.title = element_blank()) + 
  labs(x = "", y =  "") + scale_fill_viridis(discrete = T, option = "E") +
  guides(fill = guide_legend(nrow = 3))  +
  coord_cartesian(xlim = c(2020, 2055), clip = "off")
```

## Transport (Mobility)
```{r}
#| label: fig-RegionalTransportDemand 
#| fig-cap: Demand for energy (mobility transport)
# Change the working directory
plot = data |> dplyr::filter(sector == "mobilityPassenger") |> 
  group_by(year, region) |> 
  summarise(demand = sum(value, na.rm = T)) |> 
  dplyr::filter( year > 2015) |> 
  mutate(region = fct_reorder(factor(region), demand, .desc = F))
# Plot the power demand
ggplot(plot) +
  geom_bar(aes(x = year, y = demand, fill = region), alpha = 0.95, stat = "identity") +
  theme_economist() %+replace% theme(legend.title = element_blank()) + 
  labs(x = "", y =  "") + scale_fill_viridis(discrete = T, option = "E") +
  guides(fill = guide_legend(nrow = 3))
```

## HT Heat
```{r}
#| label: fig-RegionalHeatingDemand
#| fig-cap: Demand for energy (HighTempHeat)
# Change the working directory
plot = data |> dplyr::filter(sector == "heatHigh") |> 
  group_by(year, region) |> 
  summarise(demand = sum(value, na.rm = T)) |> 
  dplyr::filter( year > 2015) |> 
  mutate(region = fct_reorder(factor(region), demand, .desc = F))
# Plot the power demand
ggplot(plot) +
  geom_bar(aes(x = year, y = demand, fill = region), alpha = 0.95, stat = "identity") +
  theme_economist() %+replace% theme(legend.title = element_blank()) + 
  labs(x = "", y =  "") + scale_fill_viridis(discrete = T, option = "E") +
  guides(fill = guide_legend(nrow = 3))
```

:::


::: panel-tabset
## Dynamic Power Demand
```{r}
#| label: fig-RegionalQuarterDemand 
#| fig-cap: Demand for energy across sectors in PetaJoules
# Change the working directory
plot = data |> dplyr::filter(sector == "power", year == 2020)  |> 
  group_by(TimeStep, region) |> 
  summarise(demand = sum(value, na.rm = T)) |> 
  mutate(region = fct_reorder(factor(region), demand, .desc = F))
# Plot the power demand
ggplot(plot) +
  geom_bar(aes(x = TimeStep, y = demand, fill = region), stat = "identity") +
  theme_economist() %+replace% theme(legend.title = element_blank()) + 
  labs(x = "", y =  "") + scale_fill_viridis(discrete = T, option = "E") +
  guides(fill = guide_legend(nrow = 3))
```
## Dynamic transport (Mob) Demand
```{r}
#| label: fig-DynamicTransDemand 
#| fig-cap: Dynamic energy demand for transport (mobility)
# Change the working directory
plot = data |> dplyr::filter(sector == "mobilityPassenger", year == 2020)  |> 
  group_by(TimeStep, region) |> 
  summarise(demand = sum(value, na.rm = T)) |> 
  mutate(region = fct_reorder(factor(region), demand, .desc = F))
# Plot the power demand
ggplot(plot) +
  geom_bar(aes(x = TimeStep, y = demand, fill = region), stat = "identity") +
  theme_economist() %+replace% theme(legend.title = element_blank()) + 
  labs(x = "", y =  "") + scale_fill_viridis(discrete = T, option = "E") +
  guides(fill = guide_legend(nrow = 3))
```
## Dynamic HT Heat Demand
```{r}
#| label: fig-DynamicHeatDemand 
#| fig-cap: Dynamic energy demand for high temperature heating
# Change the working directory
plot = data |> dplyr::filter(sector == "heatHigh", year == 2020)  |> 
  group_by(TimeStep, region) |> 
  summarise(demand = sum(value, na.rm = T)) |> 
  mutate(region = fct_reorder(factor(region), demand, .desc = F))
# Plot the power demand
ggplot(plot) +
  geom_bar(aes(x = TimeStep, y = demand, fill = region), stat = "identity") +
  theme_economist() %+replace% theme(legend.title = element_blank()) + 
  labs(x = "", y =  "") + scale_fill_viridis(discrete = T, option = "E") +
  guides(fill = guide_legend(nrow = 3))
```
:::